---
title: "Project 1"
date: "15 May 2020"
output: html_document
---
## Loading and preprocessing the data

```{r, echo=TRUE}

library(readr)
activity <- read_csv("activity.csv")

```

## What is mean total number of steps taken per day?

We first ignore the missing values in the dataset.

```{r, echo=TRUE}

dailysteps <- aggregate(steps ~ date, activity , sum, na.rm=TRUE)
hist(dailysteps$steps, xlab="Daily Steps", main="Histogram for the Daily Number of Steps")

mean(dailysteps$steps)
median(dailysteps$steps)

```

The mean number of daily steps is 10766.19; the median number of daily steps is 10765.

## What is the average daily activity pattern?

```{r, echo=TRUE}

intervalsteps <- aggregate(steps ~ interval, activity, mean, na.rm=TRUE)
plot(steps~interval, intervalsteps,type="l", main="Average Daily Activity Pattern")

intervalsteps[which.max(intervalsteps$steps),]$interval

```

The interval with the highest number of average steps is 835. 

## Imputing missing values

There are 2304 rows with missing values. We impute the mean for the interval where there are missing values in the dataset. 

```{r, echo=TRUE}
sum(is.na(activity))

getMeanStepsPerInterval<-function(interval){
    intervalsteps[intervalsteps$interval==interval,]$steps
}

activityNoNA<-activity
for(i in 1:nrow(activityNoNA)){
    if(is.na(activityNoNA[i,]$steps)){
        activityNoNA[i,]$steps <- getMeanStepsPerInterval(activityNoNA[i,]$interval)
    }
}

totalStepsPerDayNoNA <- aggregate(steps ~ date, data=activityNoNA, sum)
hist(totalStepsPerDayNoNA$steps, main="Histogram after Imputation", xlab="Total Steps Per Day")

mean(totalStepsPerDayNoNA$steps)
median(totalStepsPerDayNoNA$steps)

```

After imputation, the mean number of daily steps is 10766.19; the median number of daily steps is 10765. The values do not change, so using mean imputation does not affect the mean or median. 

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
activityNoNA$date <- as.Date(strptime(activityNoNA$date, format="%Y-%m-%d"))
activityNoNA$day <- weekdays(activityNoNA$date)
for (i in 1:nrow(activityNoNA)) {
    if (activityNoNA[i,]$day %in% c("Saturday","Sunday")) {
        activityNoNA[i,]$day<-"weekend"
    }
    else{
        activityNoNA[i,]$day<-"weekday"
    }
}
stepsByDay <- aggregate(activityNoNA$steps ~ activityNoNA$interval + activityNoNA$day, activityNoNA, mean)

names(stepsByDay) <- c("interval", "day", "steps")
library(lattice)
xyplot(steps ~ interval | day, stepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")


```




